---
title: "DA_14"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(cowplot)
library(deSolve)
library(rstan)
library(coda)
library(modelbased)
library(shinystan)
library(msm)
```

//////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////
////////////////////////////////////////////////////////// DATA READ-IN AND FORMATING ////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////

READ IN RAW DATA

```{r}
#read in Data
setwd("~/Trophic_Cascade_Warming_Submission")

TCTempDataFull <- read.csv("TC_Temp_Data_Full.csv")

```

FORMAT DATES AND CALCULATE CELL DENSITIES

```{r}
#DATES
TCTempDataFull$Date <- as.Date(TCTempDataFull$Date, format = "%m/%d/%Y")

#CREATE DENSITY
TCTempDataFull <- TCTempDataFull %>% mutate(SppPerML = Count/Volume..mL.)
```

CREATE DAYS SINCE START VARIABLE

```{r}
#add days since start instead of dates
#Whoo! Thanks for the help AJ!

jars <- unique(TCTempDataFull$Jar.ID)#get a list of all the unique jars

TCTempDataFull$days.since.start <- NA#just making an empty column

min.date <- vector()#empty vector for all the minimum jar dates

for(i in jars){#iteratively figure out the min date for each jar
  dates <- TCTempDataFull$Date[TCTempDataFull$Jar.ID==i]
  min.date[i] <- min(dates)
}

TCTempDataFull$days.since.start <- TCTempDataFull$Date-min.date[TCTempDataFull$Jar.ID]#subtract!
TCTempDataFull$days.since.start <- as.numeric(TCTempDataFull$days.since.start)#fix
```

FILTER TO TARGET TEMP AND REMOVE HYDRA JARS FROM DATA

```{r}
#remove hydra dishes
TCTempDataJar_Temp_i <- TCTempDataFull %>% 
  filter(Temp == 14, Community.ID != "H-P-A")
```

FORMAT DATA INTO MATRIX FOR ODE SOLVER

```{r}
jars <- unique(TCTempDataJar_Temp_i$Jar.ID) #unique identifier for wach dish


Algae_mat <- matrix(nrow = 5, ncol = 16) #create empty matrix for algae data

for(i in 1:5) {
  Algae_mat[i,] <- TCTempDataJar_Temp_i %>% filter(Jar.ID == jars[i] & Species == 'Ank') %>% pull(SppPerML)  
} #populate empty algae matrix

Daph_mat <- matrix(nrow = 5, ncol = 16) #create empty matrix for daphnia data

for(i in 1:5) {
  Daph_mat[i,] <- TCTempDataJar_Temp_i %>% filter(Jar.ID == jars[i] & Species == 'Daph') %>% pull(SppPerML)  
} # populate empty daphnia matrix

```

///////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////
//////////////////////////////////////////////////////////////// STAN ODE FITTING /////////////////////////////////////////////////////////////
///////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////

SET STAN FITTING OPTIONS
```{r}
### set stan options

options(mc.cores = parallel::detectCores()) #identify cores in machine
rstan_options(auto_write = TRUE)
chains = 3 #declare number of MCMC chains
iter = 10000 #declare total number of fitting iterations
warmup = 3000 #declare number of iterations for burnin/warmup
```

DEFINE ODE MODEL

```{r}
# parameters: p[1] = r, p[2] = K, p[3] = a, p[4] = e, p[5] = d, p[6] = h, p[7] = w, p[8] = B
# N[1] is the Algae, N[2] is the Daphnia

odemodel <- function(t, N, p) {
  dN1dt = (p[1]*N[1]*(1-N[1]*p[2]*exp(p[8]*t)))-(p[3]*N[1]*N[2]/(1+p[7]*(N[2]-0.003333333)+p[3]*p[6]*N[1])) #equation for resource
  dN2dt = (p[4]*p[3]*N[1]*N[2]/(1+p[7]*(N[2]-0.003333333)+p[3]*p[6]*N[1]))-p[5]*N[2] #equation for consumer
  return(list(c(dN1dt, dN2dt)))
}
```

DEFINE STAN DATA

```{r}
stan_data <- list(n = length(unique(TCTempDataJar_Temp_i$days.since.start)), # n is our number of observations
                  m=length(jars),
                  t = unique(TCTempDataJar_Temp_i$days.since.start),
                  
                  N1 = Algae_mat, # Matrix of algal densities across treamtemt
                  N2 = Daph_mat #matrix of daph densities
                  )

```

IDENTIFY STAN MODEL FILE

```{r}
stanmodel <- stan_model('DA_14.stan')
```

DECLARE INITIAL VALUES FOR FITTING

```{r}
inits <- function() {
  list(r = 4.636788e-02, 
       Q =  2.858949e-07 ,
       a1 =  1.866689e-01,
       e1 =  5.279091e-04 ,
       d1 =  2.447835e+00,
       h1=   1.821810e-04,
       w1 = 2.456729e+00,
       B =  1.652951e-02,
       N1_0 = c(2.424825e+05, 2.380174e+05, 1.681849e+05, 2.366568e+05, 1.788969e+05),
       N2_0 = c(0.06666667, 0.06666667, 0.06666667, 0.06666667, 0.06666667),
       sd_obs1 =  2.571404e-01, 
       sd_obs2 = 2.571328e-01
       )
}
```

RUN STAN FIT

```{r}
DA_14_fit = sampling(stanmodel,
               data = stan_data,
               iter = iter,
               warmup = warmup,
               chains = chains,
               init = inits,
               control = list(adapt_delta =.9, max_treedepth=12.5, stepsize=1),
               refresh = 10) 
```

```{r}
summary(DA_14_fit) #FIT SUMMARY
```

DIAGNOSTIC PLOTS FOR STAN FIT

```{r}
rstan::traceplot(DA_14_fit, pars = "r", inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = "Q", inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = "a1", inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = "e1", inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = "d1", inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = "h1", inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = "w1", inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = "B", inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = 'sd_obs1', inc_warmup = TRUE)

rstan::traceplot(DA_14_fit, pars = 'sd_obs2', inc_warmup = TRUE)
```

```{r}
r_prior <- data.frame(x = seq(0,2, by = 0.01), vals =
  dtnorm(x= seq(0,2, by = 0.01), mean =4.802530e-02, sd = 9.223840e-03, lower = 0, upper = 2))

Q_prior <- data.frame(x = seq(1e-8,1e-5, by = 1e-8), vals =
  dtnorm(x= seq(1e-8,1e-5, by = 1e-8), mean = 3.541385e-07, sd =  2.720056e-07,, lower = 1e-8, upper = 1e-3))

a1_prior <- data.frame(x = seq(0,1, by = 0.001), vals =
  dtnorm(x= seq(0,1, by = 0.001), mean =2.026618e-01, sd = , 1.194885e-01, lower = 0, upper = 1))

e1_prior <- data.frame(x = seq(0,0.01, by = 1e-6), vals =
  dtnorm(x= seq(0,0.01, by = 1e-6), mean = 5.617132e-04,  sd = 2.862119e-04, lower = 0, upper = 0.01))

d1_prior <- data.frame(x = seq(0,6, by = 0.001), vals =
  dtnorm(x= seq(0,6, by = 0.001), mean = 2.380208e+00, sd = , 6.504642e-01, lower = 0))

h1_prior <- data.frame(x = seq(0,1e-3, by = 1e-6), vals =
  dtnorm(x= seq(0,1e-3, by = 1e-6), mean = 1.908917e-04, sd = 9.561157e-05, lower = 0, upper = 1e-3))

w1_prior <- data.frame(x = seq(0,4, by = 0.001), vals =
  dtnorm(x= seq(0,4, by = 0.001), mean = 2.463522e+00, sd =  3.527207e-01, lower = 0, upper = 4))

B_prior <- data.frame(x = seq(0,0.4, by = 0.0001), vals =
  dtnorm(x= seq(0,0.4, by = 0.0001), mean = 0.1, sd = 1, lower = 0, upper = 0.4))

```

```{r}
stan_dens(DA_14_fit, pars = c("r"), separate_chains = TRUE) +
   geom_line(data = r_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_14_fit, pars = c("Q"), separate_chains = TRUE) +
    geom_line(data = Q_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_14_fit, pars = c("a1"), separate_chains = TRUE) +
    geom_line(data = a1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_14_fit, pars = c("e1"), separate_chains = TRUE) +
  geom_line(data = e1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_14_fit, pars = c("d1"), separate_chains = TRUE)+
    geom_line(data = d1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_14_fit, pars = c("h1"), separate_chains = TRUE)+
    geom_line(data = h1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_14_fit, pars = c("w1"), separate_chains = TRUE) +
    geom_line(data = w1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_14_fit, pars = c("B"), separate_chains = TRUE) +
    geom_line(data = B_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_14_fit, pars = c("sd_obs1"), separate_chains = TRUE)

stan_dens(DA_14_fit, pars = c("sd_obs2"), separate_chains = TRUE)
```

```{r}
pairs(DA_14_fit, pars = c("r", "Q", "a1", "e1", "d1", "h1", "w1", "B"), log = FALSE)
```

SAVE STAN FIT AND DRAWS

```{r}
save(DA_14_fit, file = "DA_14_fit") # SAVE FIT
```

SAVE DRAWS PER ITERATION

```{r}

DA_14_draws_iter <- extract(DA_14_fit, pars = c('r', 'Q', 'a1', 'e1', 'd1', 'h1', 'w1', 'B', 'N1_0', 'N2_0'), inc_warmup = FALSE, permuted = TRUE)

DA_14_draws_iter <- as.data.frame(DA_14_draws_iter)

DA_14_draws_iter <- tibble::rownames_to_column(DA_14_draws_iter, "Iteration")

write_csv(DA_14_draws_iter, file = "DA_14_draws_iter.csv")
```

SAVE PARAMETER SUMMARY 

```{r}

DA_14_pars_sum <- extract(DA_14_fit, pars = c('r', 'Q', 'a1', 'e1', 'd1', 'h1', 'w1', 'B', 'N1_0', 'N2_0'), inc_warmup = FALSE, permuted = FALSE)

DA_14_pars_sum <- monitor(DA_14_pars_sum, warmup = 0, print = FALSE, digits_summary = 7)

DA_14_pars_sum <- as.data.frame(DA_14_pars_sum)

DA_14_pars_sum <- tibble::rownames_to_column(DA_14_pars_sum, "Parameter")

write_csv(DA_14_pars_sum, file = "DA_14_draws_sum.csv")

```



