---
title: "HDA_26"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(cowplot)
library(deSolve)
library(rstan)
library(coda)
library(modelbased)
library(shinystan)
library(msm)
```

READ IN RAW DATA

```{r}
#read in Data
setwd("/Users/francisbiagioli/Documents/R/TC Temp Experiment")

TCTempDataFull <- read.csv("TC_Temp_Data_Full.csv")

```

FORMAT DATES AND CALCULATE CELL DENSITIES

```{r}
#DATES
TCTempDataFull$Date <- as.Date(TCTempDataFull$Date, format = "%m/%d/%Y")

#CREATE DENSITY
TCTempDataFull <- TCTempDataFull %>% mutate(SppPerML = Count/Volume..mL.)
```

CREATE DAYS SINCE START VARIABLE

```{r}
jars <- unique(TCTempDataFull$Jar.ID)#get a list of all the unique jars

TCTempDataFull$days.since.start <- NA#just making an empty column

min.date <- vector()#empty vector for all the minimum jar dates

for(i in jars){#iteratively figure out the min date for each jar
  dates <- TCTempDataFull$Date[TCTempDataFull$Jar.ID==i]
  min.date[i] <- min(dates)
}

TCTempDataFull$days.since.start <- TCTempDataFull$Date-min.date[TCTempDataFull$Jar.ID]#subtract!
TCTempDataFull$days.since.start <- as.numeric(TCTempDataFull$days.since.start)#fix
```

FILTER TO TARGET TEMP AND REMOVE HYDRA JARS FROM DATA

```{r}
#remove hydra dishes
TCTempDataJar_Temp_i <- TCTempDataFull %>% 
  filter(Temp == 26, Community.ID != "P-A")
```

FORMAT DATA INTO MATRIX FOR ODE SOLVER

```{r}
jars <- unique(TCTempDataJar_Temp_i$Jar.ID) #unique identifier for wach dish


Algae_mat <- matrix(nrow = 5, ncol = 16) #create empty matrix for algae data 

for(i in 1:5) {
  Algae_mat[i,] <- TCTempDataJar_Temp_i %>% filter(Jar.ID == jars[i] & Species == 'Ank') %>% pull(SppPerML)  
} #populate empty algae matrix

Daph_mat <- matrix(nrow = 5, ncol = 16) #create empty matrix for daphnia data

for(i in 1:5) {
  Daph_mat[i,] <- TCTempDataJar_Temp_i %>% filter(Jar.ID == jars[i] & Species == 'Daph') %>% pull(SppPerML)  
} # populate empty daphnia matrix

Hydra_mat <- matrix(nrow = 5, ncol = 16) #create empty matrix for daphnia data

for(i in 1:5) {
  Hydra_mat[i,] <- TCTempDataJar_Temp_i %>% filter(Jar.ID == jars[i] & Species == 'Hydra') %>% pull(SppPerML)  
} # populate empty daphnia matrix


```

SET STAN FITTING OPTIONS
```{r}
### set stan options

options(mc.cores = parallel::detectCores()) #identify cores in machine
rstan_options(auto_write = TRUE)
chains = 3 #declare number of MCMC chains
iter = 5000 #declare total number of fitting iterations
warmup = 3000 #declare number of iterations for burnin/warmup
```

DEFINE ODE MODEL

```{r}
# parameters: p[1] = r, p[2] = K, p[3] = a, p[4] = e, p[5] = d, p[6] = h, p[7] = w, p[8] = B
# N[1] is the Algae, N[2] is the Daphnia, N[3] is Hydra

odemodel <- function(t, N, p) {
  dN1dt = (p[1]*N[1]*(1-N[1]*p[2]*exp(p[8]*t)))-(p[3]*N[1]*N[2]/(1+p[7]*(N[2]-0.003333333)+p[3]*p[6]*N[1]))
  
  dN2dt = ((p[4]*p[3]*N[1]*N[2])/((1+p[7]*(N[2]-0.003333333))+(p[3]*p[6]*N[1])))-((p[9]*N[2]*N[3])/((1+p[13]*(N[3]-0.003333333))+(p[9]*p[12]*N[2])))-(p[5]*N[2])
  
  dN3dt = ((p[10]*p[9]*N[2]*N[3])/((1+p[13]*(N[3]-0.003333333))+(p[9]*p[12]*N[2])))-(p[11]*N[3])
  
  return(list(c(dN1dt, dN2dt, dN3dt)))
}
```

DEFINE STAN DATA

```{r}
stan_data <- list(n = length(unique(TCTempDataJar_Temp_i$days.since.start)), # n is our number of observations
                  m=length(jars),
                  t = unique(TCTempDataJar_Temp_i$days.since.start),
                  
                  N1 = Algae_mat, # Matrix of algal densities across treamtemt
                  N2 = Daph_mat, #matrix of daph densities
                  N3 = Hydra_mat #matrix of hydra densities
                  )

```

IDENTIFY STAN MODEL FILE

```{r}
stanmodel <- stan_model('HDA_23.stan')
```

DECLARE INITIAL VALUES FOR FITTING

```{r}

inits <- function() {
  list(
       r = 1.520441e-01,
       Q = 1.896343e-07,
       a1 = 4.893958e+00,
       h1 = 1.088241e-05,
       w1 =  1.236402e+01,
       a2 = 3.000773e+01, 
       e2 = 8.103290e-02, 
       d2 = 3.972968e-02, 
       h2 = 1.483244e-01, 
       w2 = 8.615578e+00, 
       sd_obs1 = 2.5e5, 
       sd_obs2 = 0.2,
       sd_obs3 =  0.008,
       N1_0 = c(2e5, 2e5, 2e5, 2e5, 2e5),
       N2_0 = c(0.06666667, 0.06666667, 0.06666667, 0.06666667, 0.06666667),
       N3_0 =c(5e-3, 5e-3, 5e-3, 5e-3, 5e-3))
}

```

RUN STAN FIT

```{r}
HDA_26_Pred_Uni_fit_a1w1h1_rQ = sampling(stanmodel,
               data = stan_data,
               iter = iter,
               warmup = warmup,
               chains = chains,
               init = inits,
               control = list(adapt_delta =.8, max_treedepth=12.5, stepsize=1), refresh = 10) 
#list(adapt_delta =.9, max_treedepth=12, stepsize=0.8)
```

```{r}
summary(HDA_26_Pred_Uni_fit_a1w1h1_rQ) #FIT SUMMARY
```

DIAGNOSTIC PLOTS FOR STAN FIT

```{r}
rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "r", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "Q", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "a1", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "h1", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "w1", inc_warmup = TRUE)
```


```{r}

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "a2", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "e2", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "d2", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "h2", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = "w2", inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = 'sd_obs1', inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = 'sd_obs2', inc_warmup = TRUE)

rstan::traceplot(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = 'sd_obs3', inc_warmup = TRUE)
```

```{r}

r_prior <- data.frame(x = seq(0,1, by = 0.01), vals =
  dtnorm(x= seq(0,1, by = 0.01), mean =2.046415e-01, sd = 4.671682e-02, lower = 0))

Q_prior <- data.frame(x = seq(0,1e-6, by = 1e-9), vals =
  dtnorm(x= seq(0,1e-6, by = 1e-9), mean =2.163617e-07, sd = 9.024294e-08, lower = 0))

a1_prior <- data.frame(x = seq(0,15, by = 0.01), vals =
  dtnorm(x= seq(0,15, by = 0.01), mean =5, sd = 3, lower = 0))

h1_prior <- data.frame(x = seq(0,1e-3, by = 1e-6), vals =
  dtnorm(x= seq(0,1e-3, by = 1e-6), mean = 1.542272e-05, sd = 4.279144e-06, lower = 0))

w1_prior <- data.frame(x = seq(0,30, by = 0.01), vals =
  dtnorm(x= seq(0,30, by = 0.01), mean =8.439199e+00, sd = 6.842676e+00, lower = 0))
```

```{r}

a2_prior <- data.frame(x = seq(0,100, by = 0.01), vals =
  dtnorm(x= seq(0,100, by = 0.01), mean =3.362355e+01, sd = 1.515929e+01, lower = 0))

e2_prior <- data.frame(x = seq(0,1, by = 0.0001), vals =
  dtnorm(x= seq(0,1, by = 0.0001), mean = 8.366662e-02, sd = 2.245486e-02, lower = 0, upper = 1))

d2_prior <- data.frame(x = seq(0,1, by = 0.0001), vals =
  dtnorm(x= seq(0,1, by = 0.0001), mean = 4.068907e-02, sd = 1.157219e-02, lower = 0))

h2_prior <- data.frame(x = seq(0,1, by = 0.001), vals =
  dtnorm(x= seq(0,1, by = 0.001), mean = 1.531418e-01, sd = 7.087266e-02, lower = 0))

w2_prior <- data.frame(x = seq(0,40, by = 0.01), vals =
  dtnorm(x= seq(0,40, by = 0.01), mean = 1.203942e+01, sd = 1.174399e+01, lower = 0))


```

```{r}
stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("r"), separate_chains = TRUE) +
    geom_line(data = r_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("Q"), separate_chains = TRUE) +
    geom_line(data = Q_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("a1"), separate_chains = TRUE) +
    geom_line(data = a1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("h1"), separate_chains = TRUE) +
    geom_line(data = h1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("w1"), separate_chains = TRUE) +
    geom_line(data = w1_prior, aes(x=x, y = vals), linewidth = 0.75)

```

```{r}

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("a2"), separate_chains = TRUE) +
    geom_line(data = a2_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("e2"), separate_chains = TRUE) +
    geom_line(data = e2_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("d2"), separate_chains = TRUE) +
    geom_line(data = d2_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("h2"), separate_chains = TRUE) +
    geom_line(data = h2_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("w2"), separate_chains = TRUE) +
    geom_line(data = w2_prior, aes(x=x, y = vals), linewidth = 0.75)
```

```{r}

pairs(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c("r", "Q", "a1", "h1", "w1", "a2", "e2", "d2", "h2", "w2"), log = FALSE)

```

SAVE STAN FIT AND DRAWS

```{r}
save(HDA_26_Pred_Uni_fit_a1w1h1_rQ, file = "HDA_26_Pred_Uni_fit_a1w1h1_rQ") # SAVE FIT
```

SAVE DRAWS PER ITERATION

```{r}

HDA_26_draws_iter <- extract(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c('r', 'Q', 'a1', 'h1', 'w1', 'a2', 'e2', 'd2', 'h2', 'w2', 'N1_0', 'N2_0', 'N3_0'), inc_warmup = FALSE, permuted = TRUE)

HDA_26_draws_iter <- as.data.frame(HDA_26_draws_iter)

HDA_26_draws_iter <- tibble::rownames_to_column(HDA_26_draws_iter, "Iteration")

write_csv(HDA_26_draws_iter, file = "HDA_26_draws_iter.csv")
```

SAVE PARAMETER SUMMARY 

```{r}

HDA_26_pars_sum <- extract(HDA_26_Pred_Uni_fit_a1w1h1_rQ, pars = c('r', 'Q', 'a1', 'h1', 'w1', 'a2', 'e2', 'd2', 'h2', 'w2', 'N1_0', 'N2_0', 'N3_0'), inc_warmup = FALSE, permuted = FALSE)

HDA_26_pars_sum <- monitor(HDA_26_pars_sum, warmup = 0, print = FALSE, digits_summary = 7)

HDA_26_pars_sum <- as.data.frame(HDA_26_pars_sum)

HDA_26_pars_sum <- tibble::rownames_to_column(HDA_26_pars_sum, "Parameter")

write_csv(HDA_26_pars_sum, file = "HDA_26_pars_sum.csv")

```

