---
title: "DA_26"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(cowplot)
library(deSolve)
library(rstan)
library(coda)
library(modelbased)
library(shinystan)
library(msm)
```

READ IN RAW DATA

```{r}
#read in Data
setwd("~/Trophic_Cascade_Warming_Submission")

TCTempDataFull <- read.csv("TC_Temp_Data_Full.csv")

```

FORMAT DATES AND CALCULATE CELL DENSITIES

```{r}
#DATES
TCTempDataFull$Date <- as.Date(TCTempDataFull$Date, format = "%m/%d/%Y")

#CREATE DENSITY
TCTempDataFull <- TCTempDataFull %>% mutate(SppPerML = Count/Volume..mL.)
```

CREATE DAYS SINCE START VARIABLE

```{r}
#add days since start instead of dates

jars <- unique(TCTempDataFull$Jar.ID)#get a list of all the unique jars

TCTempDataFull$days.since.start <- NA#just making an empty column

min.date <- vector()#empty vector for all the minimum jar dates

for(i in jars){#iteratively figure out the min date for each jar
  dates <- TCTempDataFull$Date[TCTempDataFull$Jar.ID==i]
  min.date[i] <- min(dates)
}

TCTempDataFull$days.since.start <- TCTempDataFull$Date-min.date[TCTempDataFull$Jar.ID]#subtract!
TCTempDataFull$days.since.start <- as.numeric(TCTempDataFull$days.since.start)#fix
```

FILTER TO TARGET TEMP AND REMOVE HYDRA JARS FROM DATA

```{r}
#remove hydra dishes
TCTempDataJar_Temp_i <- TCTempDataFull %>% 
  filter(Temp == 26, Community.ID != "H-P-A")
```

FORMAT DATA INTO MATRIX FOR ODE SOLVER

```{r}
jars <- unique(TCTempDataJar_Temp_i$Jar.ID) #unique identifier for wach dish


Algae_mat <- matrix(nrow = 5, ncol = 16) #create empty matrix for algae data

for(i in 1:5) {
  Algae_mat[i,] <- TCTempDataJar_Temp_i %>% filter(Jar.ID == jars[i] & Species == 'Ank') %>% pull(SppPerML)  
} #populate empty algae matrix

Daph_mat <- matrix(nrow = 5, ncol = 16) #create empty matrix for daphnia data

for(i in 1:5) {
  Daph_mat[i,] <- TCTempDataJar_Temp_i %>% filter(Jar.ID == jars[i] & Species == 'Daph') %>% pull(SppPerML)  
} # populate empty daphnia matrix

```

SET STAN FITTING OPTIONS
```{r}
### set stan options

options(mc.cores = parallel::detectCores()) #identify cores in machine
rstan_options(auto_write = TRUE)
chains = 3 #declare number of MCMC chains
iter = 2000 #declare total number of fitting iterations
warmup = 1000 #declare number of iterations for burnin/warmup
```

DEFINE ODE MODEL

```{r}
# parameters: p[1] = r, p[2] = K, p[3] = a, p[4] = e, p[5] = d, p[6] = h, p[7] = w, p[8] = B
# N[1] is the Algae, N[2] is the Daphnia

odemodel <- function(t, N, p) {
  dN1dt = (p[1]*N[1]*(1-N[1]*p[2]*exp(p[8]*t)))-(p[3]*N[1]*N[2]/(1+p[7]*(N[2]-0.003333333)+p[3]*p[6]*N[1])) #equation for resource
  dN2dt = (p[4]*p[3]*N[1]*N[2]/(1+p[7]*(N[2]-0.003333333)+p[3]*p[6]*N[1]))-p[5]*N[2] #equation for consumer
  return(list(c(dN1dt, dN2dt)))
}
```

DEFINE STAN DATA

```{r}
stan_data <- list(n = length(unique(TCTempDataJar_Temp_i$days.since.start)), # n is our number of observations
                  m=length(jars),
                  t = unique(TCTempDataJar_Temp_i$days.since.start),
                  
                  N1 = Algae_mat, # Matrix of algal densities across treamtemt
                  N2 = Daph_mat #matrix of daph densities
                  )

```

IDENTIFY STAN MODEL FILE

```{r}
stanmodel <- stan_model('DA_26.stan')
```

DECLARE INITIAL VALUES FOR FITTING

```{r}
inits <- function() {
  list(r =  2.396576e-01,
       Q = 1.516709e-07 ,
       a1 = 5.362482e-01,
       e1 =  4.971543e-05 ,
       d1 = 9.148129e-01,
       h1=   3.455133e-06,
       w1 =  1.242475e+00,
       B = 1.021058e-01,
       N1_0 = c(2e5, 2e5, 2e5, 2e5, 2e5),
       N2_0 = c(0.06666667, 0.06666667, 0.06666667, 0.06666667, 0.06666667),
       sd_obs1 = 1.140371e+05, 
       sd_obs2 = 2.240365e-01)
}
```

RUN STAN FIT

```{r}
DA_26_fit = sampling(stanmodel,
               data = stan_data,
               iter = iter,
               warmup = warmup,
               chains = chains,
               init = inits,
               control = list(adapt_delta =.8, max_treedepth=12.5, stepsize=1),
               refresh = 10) 
```

```{r}
summary(DA_26_fit) #FIT SUMMARY
```

DIAGNOSTIC PLOTS FOR STAN FIT

```{r}
rstan::traceplot(DA_26_fit, pars = "r", inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = "Q", inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = "a1", inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = "e1", inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = "d1", inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = "h1", inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = "w1", inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = "B", inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = 'sd_obs1', inc_warmup = TRUE)

rstan::traceplot(DA_26_fit, pars = 'sd_obs2', inc_warmup = TRUE)
```

```{r}
r_prior <- data.frame(x = seq(0,2, by = 0.01), vals =
  dtnorm(x= seq(0,2, by = 0.01), mean = 0.06, sd = 1, lower = 0, upper = 2))

Q_prior <- data.frame(x = seq(1e-8,1e-3, by = 1e-6), vals =
  dtnorm(x= seq(1e-8,1e-3, by = 1e-6), mean = 1e-7, sd = 1e-6, lower = 1e-8, upper = 1e-3))

a1_prior <- data.frame(x = seq(0,2, by = 0.001), vals =
  dtnorm(x= seq(0,2, by = 0.001), mean =1, sd = 1, lower = 0, upper = 2))

e1_prior <- data.frame(x = seq(0,0.01, by = 1e-6), vals =
  dtnorm(x= seq(0,0.01, by = 1e-6), mean = 0.001, sd = 1, lower = 0, upper = 0.01))

d1_prior <- data.frame(x = seq(0,10, by = 0.01), vals =
  dtnorm(x= seq(0,10, by = 0.01), mean = 0.2, sd = 1, lower = 0, upper = 10))

h1_prior <- data.frame(x = seq(0,1e-3, by = 1e-6), vals =
  dtnorm(x= seq(0,1e-3, by = 1e-6), mean = 1e-8, sd = 1, lower = 0, upper = 1e-3))

w1_prior <- data.frame(x = seq(0,20, by = 0.01), vals =
  dtnorm(x= seq(0,20, by = 0.01), mean = 1, sd = 1, lower = 0))

B_prior <- data.frame(x = seq(0,0.4, by = 0.0001), vals =
  dtnorm(x= seq(0,0.4, by = 0.0001), mean = 0.1, sd = 0.5, lower = 0, upper = 0.4))


```

```{r}
stan_dens(DA_26_fit, pars = c("r"), separate_chains = TRUE) +
   geom_line(data = r_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_26_fit, pars = c("Q"), separate_chains = TRUE) +
    geom_line(data = Q_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_26_fit, pars = c("a1"), separate_chains = TRUE) +
    geom_line(data = a1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_26_fit, pars = c("e1"), separate_chains = TRUE) +
  geom_line(data = e1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_26_fit, pars = c("d1"), separate_chains = TRUE)+
    geom_line(data = d1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_26_fit, pars = c("h1"), separate_chains = TRUE)+
    geom_line(data = h1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_26_fit, pars = c("w1"), separate_chains = TRUE) +
    geom_line(data = w1_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_26_fit, pars = c("B"), separate_chains = TRUE) +
    geom_line(data = B_prior, aes(x=x, y = vals), linewidth = 0.75)

stan_dens(DA_26_fit, pars = c("sd_obs1"), separate_chains = TRUE)

stan_dens(DA_26_fit, pars = c("sd_obs2"), separate_chains = TRUE)
```

```{r}
pairs(DA_26_fit, pars = c("r", "Q", "a1", "e1", "d1", "h1", "w1", "B"), log = FALSE)
```

SAVE STAN FIT AND DRAWS

```{r}
#save(DA_26_fit, file = "DA_26_fit") # SAVE FIT
```

SAVE DRAWS PER ITERATION

```{r}

DA_26_draws_iter <- extract(DA_26_fit, pars = c('r', 'Q', 'a1', 'e1', 'd1', 'h1', 'w1', 'B', 'N1_0', 'N2_0'), inc_warmup = FALSE, permuted = TRUE)

DA_26_draws_iter <- as.data.frame(DA_26_draws_iter)

DA_26_draws_iter <- tibble::rownames_to_column(DA_26_draws_iter, "Iteration")

#write_csv(DA_26_draws_iter, file = "DA_26_draws_iter.csv")
```

SAVE PARAMETER SUMMARY 

```{r}

DA_26_pars_sum <- extract(DA_26_fit, pars = c('r', 'Q', 'a1', 'e1', 'd1', 'h1', 'w1', 'B', 'N1_0', 'N2_0'), inc_warmup = FALSE, permuted = FALSE)

DA_26_pars_sum <- monitor(DA_26_pars_sum, warmup = 0, print = FALSE, digits_summary = 7)

DA_26_pars_sum <- as.data.frame(DA_26_pars_sum)

DA_26_pars_sum <- tibble::rownames_to_column(DA_26_pars_sum, "Parameter")

#write_csv(DA_26_pars_sum, file = "DA_26_pars_sum.csv")

```

```{r}
DA_26_draws <- extract(DA_26_fit, pars = c('r', 'Q', 'a1', 'e1', 'd1', 'h1', 'w1', 'B', 'N1_0', 'N2_0'), inc_warmup = FALSE, permuted = FALSE)

DA_26_vals <- monitor(DA_26_draws, warmup = 0, print = FALSE, digits_summary = 7)

DA_26_vals <- as.data.frame(DA_26_vals)

DA_26_vals <- tibble::rownames_to_column(DA_26_vals, "Parameter")

DA_26_vals <- as.data.frame(DA_26_vals)
```

```{r}
DA_26_N1_0 <- DA_26_pars_sum %>%
   filter(Parameter %in% c("N1_0[1]", "N1_0[2]", "N1_0[3]", "N1_0[4]", "N1_0[5]"))

DA_26_N1_0_avg <- mean(DA_26_N1_0$Q50)

DA_26_N2_0 <- DA_26_pars_sum %>%
   filter(Parameter %in% c("N2_0[1]", "N2_0[2]", "N2_0[3]", "N2_0[4]", "N2_0[5]"))

DA_26_N2_0_avg <- mean(DA_26_N2_0$Q50)

```

SOLVE ODE USING MEDIAN FROM POSTERIORS

```{r}
DA_26_sol <- as.data.frame(lsoda(y =c(DA_26_N1_0_avg, 
                                       DA_26_N2_0_avg),
                             times = seq(0,35, by = 0.01),
                             func = odemodel,
                             parms = c((DA_26_pars_sum %>%
                                        filter(Parameter == "r"))$`50`,
                                       (DA_26_pars_sum %>%
                                        filter(Parameter == "Q"))$`50`,
                                       (DA_26_pars_sum %>%
                                        filter(Parameter == "a1"))$`50`,
                                       (DA_26_pars_sum %>%
                                        filter(Parameter == "e1"))$`50`,
                                       (DA_26_pars_sum %>%
                                        filter(Parameter == "d1"))$`50`,
                                       (DA_26_pars_sum %>%
                                        filter(Parameter == "h1"))$`Q50`,
                                       (DA_26_pars_sum %>%
                                        filter(Parameter == "w1"))$`50`,
                                       (DA_26_pars_sum %>%
                                        filter(Parameter == "B"))$`50`)))

```


CREATE ODE SOLUTION DATA FRAME

```{r}
  Daph_median_sol_dat <- data.frame(days.since.start = DA_26_sol[,1], SppPerML = DA_26_sol[,3])
  
  Alg_median_sol_dat <- data.frame(days.since.start = DA_26_sol[,1], SppPerML = DA_26_sol[,2])
```

CREATE BASE PLOTS FOR DYNAMICS PLOT

```{r}

base_Daph <- ggplot(data = filter(TCTempDataJar_Temp_i, Species == 'Daph'), aes(x = days.since.start, y = SppPerML, color = as.factor(Jar.ID))) + geom_point(size = 3, alpha= 0.7) +
  xlim(0, 35) +
  ylim(0, 2)


base_alg <- ggplot(data = filter(TCTempDataJar_Temp_i, Species == 'Ank'), aes(x = days.since.start, y = SppPerML, color = as.factor(Jar.ID))) + geom_point(size = 3, alpha =0.7)+
  xlim(0, 35) +
  ylim(0, 1.25e6)

```

PLOT ODE SOLUTIONS ON BASE PLOTS

```{r}
daph_plot <- base_Daph + 
    geom_line(data = Daph_median_sol_dat, aes(x= days.since.start, y = SppPerML), alpha = .9, linewidth = 1.5, inherit.aes =FALSE)

alg_plot <- base_alg + 
    geom_line(data = Alg_median_sol_dat, aes(x = days.since.start, y = SppPerML), alpha = .9, linewidth = 1.5, inherit.aes =FALSE) 
```

```{r}
daph_plot <- daph_plot + theme_cowplot() + ggtitle(label = 'Ceriodaphnia - 26 Degrees') + 
  xlab('Day') + ylab('Density (per mL)')

alg_plot <- alg_plot + theme_cowplot() + ggtitle(label = 'Ank - 26 Degrees') + 
  xlab('Day') + ylab('Density (per mL)')


daph_plot

alg_plot
```






